---
title: "DataCamp_R_IntroData"
output: html_notebook
---
# Load data
```{r}
library(openintro)
data(email50)
```

# View its structure
```{r}
str(email50)
glimpse(email50)
```
# Subset of emails with big numbers: email50_big
```{r}
email50_big <- email50  %>%
  filter(number == "big")
```
# Glimpse the subset
```{r}
glimpse(email50_big)
```

# Table of number variable
```{r}
table(email50_big$number)
```
# Drop levels
```{r}
email50_big$number <- droplevels(email50_big$number)
```
# Another table of number variable
```{r}
table(email50_big$number)
```

# Calculate median number of characters: med_num_char
```{r}
med_num_char <- median(email50$num_char)
```
# Create num_char_cat variable in email50
```{r}
email50 <- email50 %>%
  mutate(num_char_cat = ifelse(num_char < med_num_char, "below median", "at or above median"))
```  
# Count emails in each category
```{r}
table(email50$num_char_cat)
```

# Create number_yn column in email50

```{r}
email50 <- email50 %>%
  mutate(number_yn = ifelse(number == "none", "no", "yes"))
```
# Visualize number_yn
```{r}
ggplot(email50, aes(x = number_yn)) +
  geom_bar()
```

# Load ggplot2
```{r}
library(ggplot2)
```
# Scatterplot of exclaim_mess vs. num_char
```{r}
ggplot(email50, aes(x = num_char, y = exclaim_mess, color = factor(spam))) +
  geom_point()
```



